DRAFT: Harmonisation and trait aggregation paper

### 2020-12-10 Paper/Draft_harmonization_paper.tex
10.12: Discussion draft finished

\subsection*{Selection of traits and harmonization of trait databases}

@@

We selected traits of seven grouping features that were available in all databases, are commonly applied in trait-based ecological studies, and describe different parts of the biology of a species: life history (Voltinism), morphology (Respiration, Body form, Size), ecology (Locomotion, Feeding mode) and reproduction (Oviposition). We omitted ecological traits that describe habitat preferences (e.g. temperature preference) because these traits are missing in the New Zealand trait database. The grouping features were differently classified across the databases, we therefore harmonised them into 26 traits (Table \ref{tab:traits_harmonization}). Harmonisation was undertaken by amalgamating similar traits into one trait (e.g. crawlers and sprawlers into crawlers). Thereby, for a particular taxa the highest trait affinity score among the amalgamated traits was taken.

@@

\label{tab:traits_harmonization}

@@

Traits of the harmonised grouping feature datasets were aggregated to family-level using three approaches. I) Direct aggregation of taxa to family-level giving equal weight to every taxon using the mean or median, denoted \textit{direct\_agg\textsubscript{mean}} and \textit{direct\_agg\textsubscript{median}}, respectively. II) Stepwise aggregation, i.e. first to the genus-level and subsequently to the family-level using the mean or median. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{mean}} or \textit{stepwise\_agg\textsubscript{median}}, respectively. III) Aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights the genera according to the number of their species in the trait datasets regardless if for every used grouping feature information was available (Figure \ref{fig:data_proc_overview}).

@@

\subsection*{Analysis of the effect of harmonization and trait aggregation on trait-environment relationships}

We repeated the analysis of Szöcs et al. \cite{szocs_effects_2014} who studied the effect of anthropogenic salinization on invertebrates in the River Werra in Germany. For the re-analysis we used the established harmonised grouping features for Europe and additionally aggregated traits using the aforementioned aggregation methods.

@@

We substituted 6 of the grouping features from the original data with harmonised grouping features from the European trait dataset. We compared the results of the redundancy analysis (RDA) from the original study to the case when using harmonised grouping features. Specifically, the trait composition expressed as community weighted mean (CWM) traits was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA, i.e. the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots. Following the original study, we identified traits associated with high or low salinity based on their distance to the ordination axis median using the mahalanobis distance. Traits with a mahalanobis distance greater than the $97.5 \%$ -quantile of the Chi-square distribution (5.02) were regarded as traits responding to either low or high salinity. For our analysis, we used the same 21 grouping features that Szöcs et al. \cite{szocs_effects_2014} used. The 6 harmonised grouping features used were \textit{Size, Feeding mode, Locomotion, Oviposition, Respiration}, and \textit{Voltinism}. Additionally, for testing the effect of aggregated traits we assigned to each taxon in Szöcs et al. \cite{szocs_effects_2014} the aggregated trait value for its corresponding family and repeated the RDA.

@@

The data processing and aforementioned analyses were carried out using R (Version 3.6.1). Raw data and the R code for data processing and grouping feature harmonization is located in the Github repository: \url{https://github.com/KunzstLD/Invertebrate_traits}. Scripts and data to reproduce the trait aggregation and analysis with aggregated traits are located in the Github repository \url{https://github.com/KunzstLD/Trait-aggregation}.

@@

The trait aggregation methods \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}} and \textit{weighted\_agg} yielded similar ranges of aggregated trait affinities within each level of trait variability for the \textit{sim\_base} scenario. The \textit(stepwise\_agg \textsubscript{median}) and \textit(direct\_agg \textsubscript{median}) yielded mostly greater ranges of aggregated trait affinities for the \textit{sim\_base} scenario. As expected with increasing trait variability the produced range of trait affinities increased (Figure \ref{fig:overview_sim_results}). By contrast, the ranges of trait affinities slightly differed for all aggregation methods for the simulation scenarios \textit{sim\_extreme} and \textit{sim\_variation}. \textit{Weighted\_agg} and \textit{stepwise\_agg\textsubscript{mean}} produced a wider range of trait affinities than the \textit{direct\_agg\textsubscript{mean}} in the \textit{sim\_extreme} scenario. For most levels of trait variability the \textit{stepwise\_agg \textsubscript{median}} resulted in the largest range of trait affinities in the \textit{sim\_extreme} and \textit{sim\_variation} scenarios.

Each trait aggregation method yielded one aggregated affinity for every simulated dataset, in total 15000 ($100$ replicates $*$ 5 levels of trait variability $* 10$ unique comparisons of trait aggregation methods $* 3$ scenarios). Overall, differences between the results of the aggregation methods for each simulated dataset increased with increasing trait variability. However, in most simulated datasets the different aggregation methods often resulted in similar trait affinities. Only 1.42 \% (213 out of 15.000) of all aggregated affinities showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure \ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

@@

The percentage of differing cases of trait affinities obtained by the trait aggregation methods compared to trait affinities originally assigned at family-level varied between 16.2 \% and 22.9 \% for the Australian dataset. For the North American dataset, comparison of the trait aggregation methods to trait affinities assigned at family-level yielded between 15.3 \% and 47 \% differing cases (Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). In general, trait aggregation methods using the median yielded fewer cases with differences compared to approaches using the mean. However, aggregation methods using the median produced greater differences for both datasets. Standard deviations of absolute differences were similar for all tested aggregation methods. For both datasets maximum differences of 1 occurred for all investigated grouping features (Figure \ref{fig:diff_aggr_traits_chessman} and Figure \ref{fig:diff_aggr_traits_pyne}).

% Is that necessary?
A comparison of the aggregation methods with each other for the 4 datasets revealed that differences in aggregated trait affinities were largest between the \textit{stepwise\_agg\textsubscript{median}} and \textit{direct\_agg\textsubscript{median}} (Figure \ref{fig:comp_aggr_methods}).

\begin{table}[H]
  \centering
  \caption{Amount of differing cases, the minimum and maximum, and means and standard deviations of absolute differences between trait affinities assigned at family-level and aggregated trait affinities.}
  \label{tab:summary_stat_aggr_vs_fam_assigned}
  \begin{tabular}{ll|ccccc}
  \toprule[.1em]
  Database & \specialcell{Comparison to\\ traits at fam.-lvl.} & \specialcell{Differing \\ cases [\%]} & \specialcell{Min. \\ differences} & \specialcell{Max. \\ differences} & \specialcell{Mean abs. \\ differences} & \specialcell{SD abs. \\ differences} \\ 
  \toprule[.1em]
  \multirow{4}{*}{\specialcell{Australia \\ (Chessman)}} & \specialcell{direct\_agg (median)} & 16.53 & 0.01 & 1.00 & 0.45 & 0.27 \\ 
  & \specialcell{direct\_agg (mean)} & 23.24 & $< 0.01$ & 0.99 & 0.34 & 0.23 \\ 
  & \specialcell{stepwise\_agg (median)} & 17.90 & 0.01 & 1.00 & 0.42 & 0.26 \\ 
  & \specialcell{stepwise\_agg (mean)} & 23.24 & $< 0.01$ & 0.99 & 0.33 & 0.22 \\ 
  & \specialcell{weighted\_agg} & 23.24 & $< 0.01$ & 1.00 & 0.34 & 0.24 \\ 
  \midrule
  \multirow{4}{*}{\specialcell{North America}} & \specialcell{direct\_agg (median)} & 15.33 & 0.17 & 1.00 & 0.70 & 0.26 \\ 
  & \specialcell{direct\_agg (mean)} & 47.00 & $< 0.01$ & 1.00 & 0.30 & 0.26 \\ 
  & \specialcell{stepwise\_agg (median)} & 18.00 & 0.08 & 1.00 & 0.63 & 0.28 \\ 
  & \specialcell{stepwise\_agg (mean)} & 47.00 & $< 0.01$ & 1.00 & 0.30 & 0.27 \\ 
  & \specialcell{weighted\_agg} & 47.00 & $< 0.01$ & 1.00 & 0.31 & 0.28 \\ 
  \bottomrule
  \end{tabular}
\end{table}

@@

Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figure \ref{fig:rda_species_scores} and Figure \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition has been derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonization}). The traits shredder and gill respiration did not characterise sites with high salinisation.

@@

% 1. Allgemeiner Satz zum Thema, der auf eine offene Frage o.ä. hinweist
% 2. Was haben wir herausgefunden?
% 3. Was haben andere schon herausgefunden
% 4. Was bedeutet das für zukünftige Studien?

% 1. Trait Database 
%- Unterschiede zwischen Regionen und Trait Genauigkeit
%- Was kann man machen
%- Andere Studien zur Trait Harmonisierung usw.

@@

% Why do we need trait-based analysis, especially across regions?
Synthesizing trait information from multiple trait databases is crucial to unfold the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt of harmonising invertebrate trait information from trait databases of different regions showed that grouping feature harmonisation is labour-intensive because grouping features are differentiated differently into traits between trait databases and various codings are used to describe traits. In addition, the same traits are sometimes defined differently in the investigated databases, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database, Table \ref{tab:traits_harmonization}). Although, others have noted the lack of standardised trait terminology in freshwater ecology \cite{baird_toward_2011, brink_traits-based_2011} we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features.
%In addition we provide harmonised grouping feature trait datasets that can be used for other studies
To resolve definition discrepancies and facilitate data synthesis in the future terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative \cite{garnier_towards_2017}. The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by Baird et al. \cite{baird_toward_2011} almost a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS) providers of invertebrate trait data could connect their traits to such ontologies \cite{schneider_towards_2019}. Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results. % What is not mentioned here is how traits are measured, for which measurement protocols exist I believe. I think this is an separate issue.

% ?What's the problem with the state of data coverage
Our analysis aimed to use most of the available invertebrate trait information for different regions to establish harmonised grouping feature datasets. Although many taxa are covered in the trait databases from which we extracted our trait datasets, the availability of trait information varies strongly across grouping features and taxonomic resolution differs between databases. While trait information from Europe and New Zealand was mainly on species-level, considerable portions of trait data from North America and Australia were on genus or family-level (Table \ref{tab:tax_coverage}). Assigned traits on family-level might not reflect the real trait diversity within a taxon, for example for Chironomidae \cite{serra_synthesising_2016}. However, species identification can be complicated and time-consuming \cite{marshall_taxonomic_2006, resh_which_2008}, and therefore trait information on family-level has been widely used, e.g. in bioassessment \cite{beketov_spear_2009}.
% How would obtaining information on species level change our results?

% Include Ben's point?
% Discussion point BK: It is intersting that Australia has fewer species and genus than NOA and EU (not suprisngly since the datasets in Oz were mostly delveoped at family level) but more families than these other regions. I wonder if this indicates that Family richness is really higher in Australia. Perhaps it might be as Australia has a wider range of evolutionary histories. This would be worth considering in the discussion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 2. + 3. die aktuellen Abschnitte (obwohl sie auch etwas ähnliche Themen beinhalten - geht ja immer um was ähnliches - wie gut sind die Aggregationsmethoden - könnte man vielleicht auch zusammenfassen
% Hier: Überschrift mit genereller Aussage?

% Why is it important to investigate trait aggregation methods? And trait harmonisation?
When trait values are only available at species-level and observational data are on less precise taxonomic levels various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information on species or genus-level to a point estimate on family-level suggests a precision which is not necessarily present, especially not for traits with a high variability or if trait information on species-level is missing. Some traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares \cite{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable \cite{poff_functional_2006, wilkes_traitbased_2020}. However, if trait aggregation is necessary our study can give guidance, under which circumstances the choice of aggregation method might be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the aggregation methods \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{mean}} (median aggregation methods) compared to \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_mean} (mean aggregation methods) were often closer to the assigned traits at family-level and 2) that the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded in 47 \% of the cases a difference to the assigned traits, much higher in comparison to the median aggregation methods (between 15 and 18 \%, Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations was twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31, Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). The differences can be explained by the binary coding used in the North American trait dataset and in the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature have in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa will often result in a value of 0 or 1, while using the mean will often result in values between 0 and 1. 
% Was folgt daraus?
%Thus, using the median led for the North American dataset in many cases to agreement with the assigned traits. However, when differences were found the difference between median aggregated traits and assigned traits was often 1. 
% except for traits were binary coding was not mutually exclusive and when same number of 0 and 1

The weightings explored in this analysis should lead to different aggregated trait affinities in cases where genera within a family have different numbers of species. In fact, the simulation results indicated that for a particular family where one genus has a much larger number of species compared to the other genera, weighting the number of taxa can lead to different aggregated trait affinities, sspecially for traits with higher levels of variance (Figure \ref{fig:sim_indv_runs}. However, in the comparison of the trait aggregation methods to assigned traits the number of differing cases and the mean absolute differences in trait affinities were similar across the mean aggregation and across the median aggregation methods, which suggests a small influence of the weightings on the aggregation methods. Also, the distributions of absolute trait affinity differences to assigned traits were relatively similar for the mean aggregation methods and the median aggregation methods respectively (Figure \ref{fig:diff_aggr_traits_chessman} and Figure \ref{fig:diff_aggr_traits_pyne}). The minor impact on trait aggregation of the weightings may be explained by the fact that a considerable portion of taxa had low numbers of genera or species entries. From the taxa that were compared from the North American trait dataset, 14 \% were resolved on family-level and 62 \% were resolved on genus-level. 52 \% had 5 or less genera and 13 \% contained just one genus (Figure \ref{fig:tax_hierarchy_NOA}). In the Australian dataset, 21 \% of the taxa were resolved on family-level, 40 \% were resolved on genus level, and 39 \% were resolved on species-level. 68 \% of the taxa contain 5 or less genera and 40 \% just contained one genus (Figure \ref{fig:tax_hierarchy_AUS}). Hence, these results are expected to change when trait information is all on species level

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography

>>>>

\subsection*{Selection of traits and harmonisation of trait databases}

@@

We selected traits of seven grouping features that were available in all databases, are commonly applied in trait-based ecological studies, and describe different parts of the biology of a species: life history (Voltinism), morphology (Respiration, Body form, Size), ecology (Locomotion, Feeding mode) and reproduction (Oviposition). We omitted ecological traits that describe habitat preferences (e.g. temperature preference) because these traits are missing in the New Zealand trait database. The grouping features were differently classified across the databases, we therefore harmonised them into 26 traits (Table \ref{tab:traits_harmonisation}). Harmonisation was undertaken by amalgamating similar traits into one trait (e.g. crawlers and sprawlers into crawlers). Thereby, for a particular taxa the highest trait affinity score among the amalgamated traits was taken.

@@

\label{tab:traits_harmonisation}

@@

Traits of the harmonised grouping feature datasets were aggregated to family-level using three approaches. I) Direct aggregation of taxa to family-level giving equal weight to every taxon using the mean or median, denoted \textit{direct\_agg\textsubscript{mean}} and \textit{direct\_agg\textsubscript{median}}, respectively. II) Stepwise aggregation, i.e. first to the genus-level and subsequently to the family-level using the mean or median. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{mean}} or \textit{stepwise\_agg\textsubscript{median}}, respectively. III) Aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights the genera according to the number of their species in the trait datasets regardless if for every used grouping feature information was available (Figure \ref{fig:data_proc_overview}). When we refer to \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_agg} together we denote these methods as mean aggregation methods and when we refer to \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}} together, we denote these methods as median aggregation methods.

@@

\subsection*{Analysis of the effect of harmonisation and trait aggregation on trait-environment relationships}

We repeated the analysis of Szöcs et al. \cite{szocs_effects_2014} who studied the effect of anthropogenic salinization on invertebrates in the River Werra in Germany. For the re-analysis we used the established harmonised grouping feature dataset for Europe and additionally aggregated traits using the aforementioned aggregation methods.

@@

We substituted 6 of the grouping features from the original data with harmonised grouping features from the European trait dataset. We compared the results of the redundancy analysis (RDA) from the original study to the case when using harmonised grouping features. Specifically, the trait composition expressed as community weighted mean (CWM) traits was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA, i.e. the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots. Following the original study, we identified traits associated with high or low salinity based on their distance to the ordination axis median using the mahalanobis distance. CWM traits with a mahalanobis distance greater than the $97.5 \%$ -quantile of the Chi-square distribution (5.02) were regarded as traits responding to either low or high salinity. For our analysis, we used the same 21 grouping features that Szöcs et al. \cite{szocs_effects_2014} used. The 6 harmonised grouping features used were \textit{Size, Feeding mode, Locomotion, Oviposition, Respiration}, and \textit{Voltinism}. Additionally, for testing the effect of aggregated traits we assigned to each taxon in Szöcs et al. \cite{szocs_effects_2014} the aggregated trait value for its corresponding family and repeated the RDA.

@@

The data processing and aforementioned analyses were carried out using R (Version 3.6.1). Raw data and the R code for data processing and grouping feature harmonisation is located in the Github repository: \url{https://github.com/KunzstLD/Invertebrate_traits}. Scripts and data to reproduce the trait aggregation and analysis with aggregated traits are located in the Github repository \url{https://github.com/KunzstLD/Trait-aggregation}.

@@

As expected with increasing trait variability the produced range of trait affinities increased for all aggregation methods (Figure \ref{fig:overview_sim_results}). The mean aggregation methods yielded similar ranges of aggregated trait affinities within each level of trait variability for the \textit{sim\_base} scenario. The median aggregation methods yielded mostly greater ranges of aggregated trait affinities for the \textit{sim\_base} scenario. By contrast, the ranges of trait affinities slightly differed for all aggregation methods for the simulation scenarios \textit{sim\_extreme} and \textit{sim\_variation}. \textit{Weighted\_agg} and \textit{stepwise\_agg\textsubscript{mean}} produced a wider range of trait affinities than the \textit{direct\_agg\textsubscript{mean}} in the \textit{sim\_extreme} scenario. For most levels of trait variability the \textit{stepwise\_agg \textsubscript{median}} resulted in the largest range of trait affinities in the \textit{sim\_extreme} and \textit{sim\_variation} scenarios.

Each trait aggregation method yielded one aggregated affinity for every simulated dataset, in total 15000 comparisons ($100$ replicates $*$ 5 levels of trait variability $* 10$ unique comparisons of trait aggregation methods $* 3$ scenarios). Overall, differences between the results of the aggregation methods for each simulated dataset increased with increasing trait variability. However, in most simulated datasets the different aggregation methods often resulted in similar trait affinities. Only 1.42 \% (213 out of 15.000) of all comparisons showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure \ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

@@

The percentage of differing cases of trait affinities obtained by the trait aggregation methods compared to trait affinities originally assigned at family-level varied between 16.2 \% and 22.9 \% for the Australian dataset. For the North American dataset, comparison of the trait aggregation methods to trait affinities assigned at family-level yielded between 15.3 \% and 47 \% differing cases (Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). In general, median aggregation methods yielded fewer cases with differences compared to mean aggregation methods. However, median aggregation methods produced greater differences for both datasets. Standard deviations of absolute differences were similar for all tested aggregation methods. For both datasets maximum differences of 1 occurred for all investigated grouping features (Figure \ref{fig:diff_aggr_traits_chessman} and Figure \ref{fig:diff_aggr_traits_pyne}).

% Is that necessary?
% A comparison of the aggregation methods with each other for the 4 datasets revealed that differences in aggregated trait affinities were largest between the \textit{stepwise\_agg\textsubscript{median}} and \textit{direct\_agg\textsubscript{median}} (Figure \ref{fig:comp_aggr_methods}).

\begin{table}[H]
  \centering
  \caption{Amount of differing cases, minimum, maximum, means, and standard deviations of absolute differences between trait affinities assigned at family-level and aggregated trait affinities.}
  \label{tab:summary_stat_aggr_vs_fam_assigned}
  \begin{tabular}{ll|ccccc}
  \toprule[.1em]
  Database & \specialcell{Comparison to\\ traits at fam.-lvl.} & \specialcell{Differing \\ cases [\%]} & \specialcell{Min. \\ differences} & \specialcell{Max. \\ differences} & \specialcell{Mean abs. \\ differences} & \specialcell{SD abs. \\ differences} \\ 
  \toprule[.1em]
  \multirow{4}{*}{\specialcell{Australia \\ (Chessman)}} & \specialcell{direct\_agg (median)} & 16.53 & 0.01 & 1.00 & 0.45 & 0.27 \\ 
  & \specialcell{direct\_agg (mean)} & 23.24 & $< 0.01$ & 0.99 & 0.34 & 0.23 \\ 
  & \specialcell{stepwise\_agg (median)} & 17.90 & 0.01 & 1.00 & 0.42 & 0.26 \\ 
  & \specialcell{stepwise\_agg (mean)} & 23.24 & $< 0.01$ & 0.99 & 0.33 & 0.22 \\ 
  & \specialcell{weighted\_agg} & 23.24 & $< 0.01$ & 1.00 & 0.34 & 0.24 \\ 
  \midrule
  \multirow{4}{*}{\specialcell{North America}} & \specialcell{direct\_agg (median)} & 15.33 & 0.17 & 1.00 & 0.70 & 0.26 \\ 
  & \specialcell{direct\_agg (mean)} & 47.00 & $< 0.01$ & 1.00 & 0.30 & 0.26 \\ 
  & \specialcell{stepwise\_agg (median)} & 18.00 & 0.08 & 1.00 & 0.63 & 0.28 \\ 
  & \specialcell{stepwise\_agg (mean)} & 47.00 & $< 0.01$ & 1.00 & 0.30 & 0.27 \\ 
  & \specialcell{weighted\_agg} & 47.00 & $< 0.01$ & 1.00 & 0.31 & 0.28 \\ 
  \bottomrule
  \end{tabular}
\end{table}

@@

Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figure \ref{fig:rda_species_scores} and Figure \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition has been derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonisation}). The traits shredder and gill respiration did not characterise sites with high salinisation.

@@

@@

Synthesizing trait information from multiple trait databases is crucial to unfold the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt of harmonising invertebrate trait information from trait databases of different regions showed that grouping feature harmonisation is labour-intensive because grouping features are differentiated differently into traits between trait databases and various codings are used to describe traits. In addition, the same traits are sometimes defined differently in the investigated databases, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database, Table \ref{tab:traits_harmonisation}). Although, others have noted the lack of standardised trait terminology in freshwater ecology \cite{baird_toward_2011, brink_traits-based_2011} we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features.
%In addition we provide harmonised grouping feature trait datasets that can be used for other studies
To resolve definition discrepancies and facilitate data synthesis in the future terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative \cite{garnier_towards_2017}. The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by Baird et al. \cite{baird_toward_2011} almost a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS) providers of invertebrate trait data could connect their traits to such ontologies \cite{schneider_towards_2019}. Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results. % What is not mentioned here is how traits are measured, for which measurement protocols exist I believe. I think this is a separate issue.

Our analysis aimed to use most of the available invertebrate trait information for different regions to establish harmonised grouping feature datasets. Although many taxa are covered in the trait databases from which we extracted our trait datasets, the availability of trait information varies strongly across grouping features and taxonomic resolution differs between databases. Much information is available for grouping features that are often used in trait-environment studies, such as feeding mode and respiration. Surprisingly, there is little information on body form, a trait that is relatively easy to determine.

While trait information from Europe and New Zealand was mainly on species-level, considerable portions of trait data from North America and Australia were on genus or family-level (Table \ref{tab:tax_coverage}). Assigned traits on family-level might not reflect the real trait diversity within a taxon, for example for Chironomidae \cite{serra_synthesising_2016}. However, species identification can be complicated and time-consuming \cite{marshall_taxonomic_2006, resh_which_2008}, and therefore trait information on family-level has been widely used, e.g. in bioassessment \cite{beketov_spear_2009}.

% Include Ben's point? Discussion on evolutionary history?
% Discussion point BK: It is intersting that Australia has fewer species and genus than NOA and EU (not suprisngly since the datasets in Oz were mostly delveoped at family level) but more families than these other regions. I wonder if this indicates that Family richness is really higher in Australia. Perhaps it might be as Australia has a wider range of evolutionary histories. This would be worth considering in the discussion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Hier: Überschrift mit genereller Aussage?
\subsection*{Trait aggregation}

When trait values are only available at species-level and observational data are on less precise taxonomic levels various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information on species or genus-level to a point estimate on family-level suggests a precision which is not necessarily present, especially not for traits with a high variability or if trait information on species-level is missing. Some traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares \cite{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable \cite{poff_functional_2006, wilkes_traitbased_2020}. However, if trait aggregation is necessary our study can give guidance, under which circumstances the choice of aggregation method might be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) that the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded in 47 \% of the cases a difference to the assigned traits, much higher in comparison to the median aggregation methods (between 15 and 18 \%, Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31, Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). The differences can be explained by the binary coding used in the North American trait dataset and in the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature have in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa will often result in a value of 0 or 1, while using the mean will often result in values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

The weightings explored in this analysis should lead to different aggregated trait affinities in cases where genera within a family have different numbers of species. In fact, the simulation results indicated that for a particular family where one genus has a much larger number of species compared to the other genera, weighting the number of taxa can lead to different aggregated trait affinities, especially for traits with higher levels of variance (Figure \ref{fig:sim_indv_runs}. However, in the comparison of the trait aggregation methods to assigned traits the number of differing cases and the mean absolute differences in trait affinities were similar across the mean aggregation and across the median aggregation methods, which suggests a small influence of the weightings on the aggregation methods. Also, the distributions of absolute trait affinity differences to assigned traits were relatively similar for the mean aggregation methods and the median aggregation methods respectively (Figure \ref{fig:diff_aggr_traits_chessman} and Figure \ref{fig:diff_aggr_traits_pyne}). The minor impact on trait aggregation of the weightings may be explained by the fact that a considerable portion of taxa had low numbers of genera or species entries. From the taxa that were compared from the North American trait dataset, 14 \% were resolved on family-level and 62 \% were resolved on genus-level. 52 \% had 5 or less genera and 13 \% contained just one genus (Figure \ref{fig:tax_hierarchy_NOA}). In the Australian dataset, 21 \% of the taxa were resolved on family-level, 40 \% were resolved on genus level, and 39 \% were resolved on species-level. 68 \% of the taxa contain 5 or less genera and 40 \% just contained one genus (Figure \ref{fig:tax_hierarchy_AUS}). Hence, these results might change when more trait information on species-level becomes available.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography
